---
title: "finalproject"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(plotly)
library(RColorBrewer)
```

####Loading and tidying the Yelp business dataset

```{r tidy_bus}
business <- read_csv("./data/business.csv") %>% 
  mutate(name = str_sub(name, 2, -2)) %>% 
  mutate(address = str_sub(address, 2, -2)) %>% 
  filter(neighborhood != "Downtown Tampa" & neighborhood != "North Valley")

categories <- business %>% 
  select(business_id, categories) %>% 
  separate(categories, into = c("cat1", "cat2", "cat3", "cat4", "cat5", "cat6", "cat7", "cat8", "cat9", "cat10", "cat11", "cat12", "cat13", "cat14", "cat15", "cat16", "cat17", "cat18", "cat19", "cat20", "cat21", "cat22", "cat23", "cat24", "cat25"), sep  = ";") %>% 
  gather(key = "cat_id", value = "category", starts_with("cat")) %>% 
  select(-cat_id) 
  
restaurant_ids <- categories %>% 
  filter(category == "Restaurants") %>% 
  distinct(business_id)

restaurants <- business %>% 
  mutate(is_rest = ifelse(str_detect(categories, ";Restaurant") == TRUE, 1, 
                                ifelse(str_match(categories, ";Food") == TRUE, 1, 0))) %>% 
  filter(is_rest == 1) %>% 
  select(-is_rest) 
          
```

####Loading and tidying the attributes dataset

```{r tidy_att}
attributes <- read_csv("./data/attributes.csv") %>% 
  janitor::clean_names() %>% 
  select(business_id, alcohol)
```

####Joining the business and attributes dataset

```{r join}
restaurants <- restaurants %>% 
  left_join(attributes, by = "business_id")
```

#### exploratory stuff

```{r exploring}
restaurants %>% 
  group_by(neighborhood) %>% 
  count() %>% 
  arrange(n) %>% 
  knitr::kable()

restaurants %>% 
select(business_id, neighborhood, stars) %>% 
  distinct() %>% 
  group_by(neighborhood, stars) %>% 
  tally %>% 
  rename(my_count = n) %>% 
  spread(key = stars, value = my_count) %>% 
  knitr::kable()

restaurants %>%
  mutate(review_count = as.numeric(review_count)) %>% 
  select(business_id, neighborhood, review_count) %>% 
  group_by(neighborhood) %>% 
  summarise(Average_Number_of_Reviews = mean(review_count)) %>% 
  knitr::kable()

restaurants %>% 
  group_by(neighborhood) %>% 
  summarize(avg = mean(stars), sd = sd(stars)) %>% 
  arrange(avg) %>% 
  knitr::kable()
```


####Creating a plotly of restaurant locations

Center of Las Vegas: 36.1699° N, 115.1398° W. Plotly displayed is surrounds this center by 0.5 degrees longitude and latitude.

```{r stephs_geo_plot}
restaurants %>% 
  filter(latitude > 35.6699 & latitude < 36.6699) %>% 
  filter(longitude < -114.6398 & longitude > -115.6398) %>%
  plot_ly(x = ~longitude, y = ~latitude, type = "scatter", mode = "markers",
          alpha = 0.5, 
          color = ~stars, hoverinfo = 'text',
        text = ~paste(name, " @", neighborhood, "\n", address, "\n", city, ", ", state, postal_code, "\n", stars, "stars on Yelp")) %>%
  layout(xaxis = list(title = "Longitude"),
         yaxis = list(title = "Latitude"))
```

####Creating a plotly of open and closed restaurant compared to their rating on Yelp

```{r open_clsoe}
restaurants %>% 
  mutate(stars = if_else(stars == 1, "1",
                         if_else(stars == 1.5, "1.5",
                                 if_else(stars == 2, "2",
                                         if_else(stars == 2.5, "2.5",
                                                 if_else(stars == 3, "3",
                                                         if_else(stars == 4, "4",
                                                                 if_else(stars == 4.5, "4.5", "5"))))))),
         review_count = as.numeric(review_count)) %>% 
  group_by(stars) %>% 
  plot_ly(x = ~stars, y = ~review_count, color = ~stars, type = "bar", colors = "Set3") %>% 
  layout(xaxis = list(title = "Stars"),
         yaxis = list(title = "Number of Reviews"))
```

#### Plots of Restaurants
```{r sarah_stacked_bar}
popular <- categories %>% 
  filter(category == "Restaurants" | category == "Food") %>% 
  distinct(business_id) %>% 
  left_join(categories, by = "business_id") %>% 
  filter(category %in% c("Bars", "Breakfast & Brunch", "Chinese", "Italian", "Mexican", "Chicken Wings", "Salad", "Sushi Bars", "Pizza", "Steakhouses", "Fast Food"))

restaurants %>% 
  select(business_id, neighborhood) %>% 
  inner_join(popular) %>% 
  distinct() %>% 
  group_by(neighborhood, category) %>% 
  tally() %>% 
  plotly::plot_ly(x = ~neighborhood, y = ~n, type = 'bar', color = ~category, hoverinfo = 'text',
        text = ~paste(neighborhood, " has ",
                      n, " ", category, " restaurants.")) %>%
  layout(yaxis = list(title = "Restaurants"), xaxis = list(title = "", tickangle = -45), barmode = 'stack')
```

```{r}
restaurants %>% 
  inner_join(popular) %>% 
  select(stars, category) %>% 
  group_by(category) %>% 
  summarise(avg_stars = mean(stars)) %>% 
  arrange(desc(avg_stars)) %>% 
  knitr::kable()
```


#### Geographic Plot by Categories
```{r sarah_geo_plot}
restaurants %>% 
  inner_join(popular) %>% 
  plot_ly(x = ~longitude, y = ~latitude, type = "scatter", mode = "markers",
          alpha = 0.9, 
          color = ~category, hoverinfo = 'text',
        text = ~paste(name, " @", neighborhood, "\n", address, "\n", city, ", ", state, postal_code, "\n", stars, "star", category, "on Yelp.")) %>%
  layout(xaxis = list(title = "Longitude"),
         yaxis = list(title = "Latitude"))
```

<<<<<<< HEAD



=======
# STATS WOO!
Do star ratings vary by category type?
```{r}
aov_stars_by_category <- restaurants %>% 
  inner_join(popular) %>% 
  aov(stars ~ category, data = .) %>% 
  broom::glance()
```

Do stars vary by neighborhood?
```{r}
aov_stars_by_neighborhood <- restaurants %>% 
  inner_join(popular) %>% 
  aov(stars ~ neighborhood, data = .) %>% 
  broom::glance()
```

Are stars and review counts correlated?
```{r}
rsq_stars_review_count <- restaurants %>% 
  select(review_count, stars) %>% 
  lm(stars ~ review_count, data = .) %>% 
  broom::glance()
```
>>>>>>> 33f12023d833321d738b1906b34cb4b704c4914e


