---
title: "Report"
---
#####Group Members 
Sarah Vititoe (sev2125)
Stephanie Izard (smi2112)
Anusheh Kafi (ak4123)

#####Motivation: 
######Provide an overview of the project goals and motivation.
The motivation of this project is to use the Yelp dataset to look for trends in highly rated restaurants in Las Vegas. Our goal is to use the data to generate relevant insights that can be used to inform restaurant goers through dynamic visualizations to help answer the age-old question, "I don't know, what are you in the mood to eat tonight?".

#####Related work: 
######Anything that inspired you, such as a paper, a web site, or something we discussed in class.
Yelp inspired us to explore data regarding restaurants as it is commonly used on a daily basis to check out local restaurants for a wide-range of information and reviews. 
We were also inspired by the map created in class using the AirBnB data that used the latitudes and longitudes, thus found a dataset that included such thing so we would be able to create a similar map for restaurants in Las Vegas. 

#####Initial questions: 
######What questions are you trying to answer? How did these questions evolve over the course of the project? What new questions did you consider in the course of your analysis?
We are trying to answer questions about restaurant information and ratings for different neighborhoods in Las Vegas. One of the things we were interested in was seeing if people were more likely to review a restaurant if they had a more extreme opinion on their experience (like a 1 star or 5 star rating) Another thing we wanted to look at was if there were any geographic areas that had a high concentration of low rated or high rated restaurants.Another area of interest was breakdown of restaurant categories by neighborhood to see if certain neighborhood had concentrations of a specific type of restaurant.

#####Data: 
######Source
The source of the data:  [Yelp](https://www.yelp.com/) is a well known website that allows users to review and rate restaurants they've been to, so that this knowledge can be shared with others. Yelp users, or "Yelpers" have written over 163 million reviews since the website's inception in 2014, with 32 million unique users every month. This dataset was released for academic use, and can be accessed [here](https://www.kaggle.com/yelp-dataset/yelp-dataset/version/6) on Kaggle.com. [Kaggle.com](https://www.kaggle.com/) is an online community of data scientists, owned by [Google](https://www.google.com/about/), and Yelp posted these csv files to Kaggle.com as a user. 
#####Cleaning
The category tags were concatenated into 1 variable so we had to split the string and then tidy the dataset to give us a list of business id's and their matching tags. We filtered the data to only include restaurants, as this dataset was for all businesses in Las Vegas. We also removed Downtown Tampa and North Valley because thy only had one restaurant, thus it seemed like an error. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(plotly)
library(RColorBrewer)
```

```{r tidy_bus, include = FALSE}
business <- read_csv("./data/business.csv") %>% 
  mutate(name = str_sub(name, 2, -2)) %>% 
  mutate(address = str_sub(address, 2, -2)) %>% 
  filter(neighborhood != "Downtown Tampa" & neighborhood != "North Valley")

categories <- business %>% 
  select(business_id, categories) %>% 
  separate(categories, into = c("cat1", "cat2", "cat3", "cat4", "cat5", "cat6", "cat7", "cat8", "cat9", "cat10", "cat11", "cat12", "cat13", "cat14", "cat15", "cat16", "cat17", "cat18", "cat19", "cat20", "cat21", "cat22", "cat23", "cat24", "cat25"), sep  = ";") %>% 
  gather(key = "cat_id", value = "category", starts_with("cat")) %>% 
  select(-cat_id) 
  
restaurant_ids <- categories %>% 
  filter(category == "Restaurants") %>% 
  distinct(business_id)

restaurants <- business %>% 
  mutate(is_rest = ifelse(str_detect(categories, ";Restaurant") == TRUE, 1, 
                                ifelse(str_match(categories, ";Food") == TRUE, 1, 0))) %>% 
  filter(is_rest == 1) %>% 
  select(-is_rest) 
          
```

```{r tidy_att, include = FALSE}
attributes <- read_csv("./data/attributes.csv") %>% 
  janitor::clean_names() %>% 
  select(business_id, alcohol)
```

```{r join, include = FALSE}
restaurants <- restaurants %>% 
  left_join(attributes, by = "business_id")

popular <- categories %>% 
  filter(category == "Restaurants" | category == "Food") %>% 
  distinct(business_id) %>% 
  left_join(categories, by = "business_id") %>% 
  filter(category %in% c("Bars", "Breakfast & Brunch", "Chinese", "Italian", "Mexican", "Chicken Wings", "Salad", "Sushi Bars", "Pizza", "Steakhouses", "Fast Food"))
```
#####Exploratory analysis: 
######Visualizations, summaries, and exploratory statistical analyses. Justify the steps you took, and show any major changes to your ideas.

Our statistical analyses includes a count of the restaurants broken down by neighborhood, as seen in this table:
```{r exploring, include = TRUE}
restaurants %>% 
  group_by(neighborhood) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  rename("Restaurant Count" = n ) %>% 
  rename(Neighborhood = neighborhood) %>% 
  knitr::kable()
```

We also created a breakdown count of stars ratings for each neighborhood:

```{r exploring 2, include = TRUE}
restaurants %>% 
select(business_id, neighborhood, stars) %>% 
  distinct() %>% 
  group_by(neighborhood, stars) %>% 
  tally %>% 
  rename(Star_Count = n) %>% 
  mutate(stars = str_c(stars, " star")) %>% 
  spread(key = stars, value = Star_Count) %>% 
  arrange(desc(`5 star`),desc(`4.5 star`)) %>% 
  knitr::kable()
```

There is also a table showing the average number of reviews for restaurants in each neighborhood below. 

```{r exploring 3, include = TRUE}
restaurants %>%
  mutate(review_count = as.numeric(review_count)) %>% 
  select(business_id, neighborhood, review_count) %>% 
  group_by(neighborhood) %>% 
  summarise("Average Number of Reviews" = round(mean(review_count),0)) %>% 
  arrange(desc(`Average Number of Reviews`)) %>% 
  rename(Neighborhood = neighborhood) %>% 
  knitr::kable()
```

The following table shows the average star rating for the restaurants in each neighborhood.

```{r exploring 4, include = TRUE}
restaurants %>% 
  group_by(neighborhood) %>% 
  summarize("Average Rating" = round(mean(stars),2), "Standard Deviation" = round(sd(stars),2)) %>% 
  arrange(desc(`Average Rating`)) %>% 
  rename(Neighborhood = neighborhood) %>% 
  knitr::kable()
```

This last table shows the star ratings for different categories of restaurant types. 

```{r exploring 5, include = TRUE}
restaurants %>% 
  inner_join(popular, by = "business_id") %>% 
  select(stars, category) %>% 
  group_by(category) %>% 
  summarise(avg_stars = round(mean(stars),2)) %>% 
  arrange(desc(avg_stars)) %>% 
  rename(Category = category) %>% 
  rename("Average Star Rating" = avg_stars) %>% 
  knitr::kable()
```

```{r open_close, include = FALSE}
restaurants %>% 
  mutate(stars = if_else(stars == 1, "1",
                         if_else(stars == 1.5, "1.5",
                                 if_else(stars == 2, "2",
                                         if_else(stars == 2.5, "2.5",
                                                 if_else(stars == 3, "3",
                                                         if_else(stars == 4, "4",
                                                                 if_else(stars == 4.5, "4.5", "5"))))))),
         review_count = as.numeric(review_count)) %>% 
  group_by(stars) %>% 
  plot_ly(x = ~stars, y = ~review_count, color = ~stars, type = "bar", colors = "Set3") %>% 
  layout(xaxis = list(title = "Stars"),
         yaxis = list(title = "Number of Reviews"))
```

```{r, include = FALSE}
rsq_stars_review_count <- restaurants %>% 
  select(review_count, stars) %>% 
  lm(stars ~ review_count, data = .) %>% 
  broom::glance()

rsq_stars_review_count %>% select(adj.r.squared) %>% mutate(adj.r.squared = round(adj.r.squared, 3))

```

We believed the number of reviews may be related to the number of stars and thus plotted the two variables. It seemed that as the number of reviews increased, the star rating also increased. Our adjusted r-squared value was `r rsq_stars_review_count %>% select(adj.r.squared) %>% mutate(adj.r.squared = round(adj.r.squared, 3))`, thus we elected to not show this plot as it was not a significant relationship. 

To visualize if there were any geographic areas that had a high concentration of low rated or high rated restaurants, we plotted longitudal against latitudinal coordinates in a scatter plot. We also did a similar graph, looking for areas of high concentration of different category types of restaurants. For the stacked bar plot, we looked at the top 15 categories of restaurants, and hand selected 10 categories likely to be mutually exclusive. This plot shows the frequency of different restaurant categories, broken down by neighborhood. This plots can be seen in our [flex dashboard](flexdashboard.html). 

#####Additional analysis: 
```{r, include = FALSE}
aov_stars_by_category <- restaurants %>% 
  inner_join(popular) %>% 
  aov(stars ~ category, data = .) %>% 
  broom::glance()
```

```{r, include = FALSE}
aov_stars_by_neighborhood <- restaurants %>% 
  inner_join(popular) %>% 
  aov(stars ~ neighborhood, data = .) %>% 
  broom::glance()
```

We were interested in statistically testing whether the category of a restaurant affected the star ratings. At alpha = 0.05, we have sufficient evidence to conclude that at least one of the category's star rating differs from the others *(ANOVA: F-statistic = `r aov_stars_by_category %>% select(statistic) %>% mutate(statistic = round(statistic, 3))`, Adjusted R-Square = `r aov_stars_by_category %>% select(adj.r.squared) %>% mutate(adj.r.squared = round(adj.r.squared, 3))`, p-value = `r aov_stars_by_category %>% select(p.value) %>% mutate(p.value = round(p.value, 3))`, degrees of freedom = `r aov_stars_by_category %>% select(df) %>% mutate(df = round(df, 3))`)*. At least one category of restaurants has a different mean star value.

We were also interested in statisticallly testing whether the neighborhood the restaurant was in affected the restaurant's star rating. At alpha = 0.05, we have sufficient evidence to conclude that at least one of the neighborhood's star ratings differs from the rest *(ANOVA: F-statistic = `r aov_stars_by_neighborhood %>% select(statistic) %>% mutate(statistic = round(statistic, 3))`, Adjusted R-Square = `r aov_stars_by_neighborhood %>% select(adj.r.squared) %>% mutate(adj.r.squared = round(adj.r.squared, 3))`, p-value = `r aov_stars_by_neighborhood %>% select(p.value) %>% mutate(p.value = round(p.value, 3))`, degrees of freedom = `r aov_stars_by_neighborhood %>% select(df) %>% mutate(df = round(df, 3))`)*. At least one neighborhood has a different mean star value than the other neighborhoods. 

#####Discussion: 
######What were your findings? Are they what you expect? What insights into the data can you make?
There are 3476 restaurants included. The Strip has the most restaurants (613) and Anthem as the least (10). 

Even though The Strip has the highest number of restaurants, it is the 3rd worst rated neighborhood in terms of restaurant rating. 

Sushi bars have the highest average star rating regardless of neighborhood (3.8 stars). On the other hand, fast food places have the lowest average star rating (2.6 stars). 

PLOT 1: Contrary to our hypothesis regarding the relationship between number of ratings and star value, we found a right skewedness in number of ratings by star value. There were more reviews for the higher star values. 

PLOT 2: There were no obvious visual trends for concentration of low rated or high rated restaurants.

PLOT 3: From this plot, we see that The Strip contains the most bars, breakfast and brunch, italian, and steakhouses compared to other resturants. 

PLOT 4: same as 3 but in map form
<<<<<<< HEAD
```{r stars_geo_plot, include = FALSE}
#Center of Las Vegas: 36.1699° N, 115.1398° W. Plotly displayed is surrounds this center by 0.5 degrees longitude and latitude.
restaurants %>% 
  filter(latitude > 35.6699 & latitude < 36.6699) %>% 
  filter(longitude < -114.6398 & longitude > -115.6398) %>%
  plot_ly(x = ~longitude, y = ~latitude, type = "scatter", mode = "markers",
          alpha = 0.5, 
          color = ~stars, hoverinfo = 'text',
        text = ~paste(name, " @", neighborhood, "\n", address, "\n", city, ", ", state, postal_code, "\n", stars, "stars on Yelp")) %>%
  layout(xaxis = list(title = "Longitude"),
         yaxis = list(title = "Latitude"))
```

```{r stacked_bar, include = FALSE}
restaurants %>% 
  select(business_id, neighborhood) %>% 
  inner_join(popular) %>% 
  distinct() %>% 
  group_by(neighborhood, category) %>% 
  tally() %>% 
  plotly::plot_ly(x = ~neighborhood, y = ~n, type = 'bar', color = ~category, hoverinfo = 'text',
        text = ~paste(neighborhood, " has ",
                      n, " ", category, " restaurants.")) %>%
  layout(yaxis = list(title = "Restaurants"), xaxis = list(title = "", tickangle = -45), barmode = 'stack')
```

```{r category_geo_plot, include = FALSE}
restaurants %>% 
  inner_join(popular) %>% 
  plot_ly(x = ~longitude, y = ~latitude, type = "scatter", mode = "markers",
          alpha = 0.9, 
          color = ~category, hoverinfo = 'text',
        text = ~paste(name, " @", neighborhood, "\n", address, "\n", city, ", ", state, postal_code, "\n", stars, "star", category, "on Yelp.")) %>%
  layout(xaxis = list(title = "Longitude"),
         yaxis = list(title = "Latitude"))
```